A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-03-10 paper/draft.tex
add additional info on wages

\subsection{Why engage in patronage?}

@@

If public sector jobs are scarce, and there are only so many spots that can be filled post-election, a different form of consideration takes place from the standpoint of the politician. Differently from an electoral race in which each voter theoretically contributes the same exact amount to the race, a single vote, constituencies can vary widely with respect to the resources that they bring in to the table. A well established literature on the influence of money on politics has documented how campaign contributions play a crucial role in electoral outcomes in both the developed and developing world \citep{claessens2008political}. What is overseen is the fact that patronage can lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.

This is in fact what the data shows.

@@

\begin{figure}[H]
    \centering
    \includegraphics{figures/maps/pooled.png}
    \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}. Darker colors indicate a larger degree of partisan affiliation. This visualization includes all bureaucrats, from high-level managers to service staff.}
    \label{fig:map_pooled}
\end{figure}

>>>>

\subsection{Locking-In Wealthy Patrons}

@@

If public sector jobs are scarce, and there are only so many spots that can be filled post-election, a different form of consideration takes place from the standpoint of the politician. Differently from an electoral race in which each voter theoretically contributes the same exact amount to the race, a single vote, constituencies can vary widely with respect to the resources that they bring in to the table. A well established literature on the influence of money on politics has documented how campaign contributions play a crucial role in electoral outcomes in both the developed and developing world \citep{claessens2008political}. What is underemphasized is the fact that patronage can lock-in wealthy patrons into the bureaucracy and incentivize greater campaign contributions in the future, effectively securing important revenue streams for both the politician and her party \citep{robinson2013political}.

This is in fact what the data suggests. First, on average party members who enter the bureaucracy are wealthier than their non-partisan counterparts, in particular at the moment that the mayor assumes office. With the RAIS data, it is possible to identify for each and every employee the last wage in the non-municipal sector prior to public sector employment. Figure \ref{fig:wage_differential} plots the median of the last non-municipal sector wage by partisanship status. The gap between partisan and non-partisan wages is the highest for the beginning of the mayoral term, which is often associated with more intense patronage activity.

\begin{figure}[H]
    \centering
    \includegraphics[width = 10cm, height = 8cm]{figures/partisanship/plot_median_wage.pdf}
    \caption{\textbf{Wage Gap between Partisans and Non-Partisans}. The graph tracks the time trend in the median wage }
    \label{fig:wage_differential}
\end{figure}

Additionally

@@

\begin{figure}[H]
    \centering
    \includegraphics[width = 10cm, height = 8cm]{figures/maps/pooled.png}
    \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}. Darker colors indicate a larger degree of partisan affiliation. This visualization includes all bureaucrats, from high-level managers to service staff.}
    \label{fig:map_pooled}
\end{figure}
